Compositional equivalences based on Open pNets

### 2021-10-12 2016-Weak-FH-Bisim/2021-09-letter2Reviewers.tex
bullet in letter 2 reviewers

We detail below the major changes we made in this revision. After
this, we will review in the next pages the comments of the reviewers
and explain how we addressed them.  \emph{Comments that could be
  addressed trivially and exactly as suggested by the reviewers are
  omitted}.  \bigskip

In this new version, you will find the major improvements listed below:
\begin{itemize}
\item 
\end{itemize}

>>>>

We detail below the major changes we made in this revision. After
this, we will review in the next pages the comments of the reviewers
and explain how we addressed them.  \emph{Comments that could be
  addressed trivially and exactly as suggested by the reviewers are
  omitted}.  \medskip

In this new version, you will find the major improvements listed below:
\begin{itemize}
\item We revised the introduction with explanations on the two formalisms and why both are useful in the context of this article; we also better explained the characteristics of FH-bisimulation, and the nature of our formalisms. We also better highlight the novelty of the results.

To summarise, in order to equip  pNets with a bisimulation relation we need to define transitions that takes holes into account and are symbolic. This is exactly the characteristics of open transitions (and open automata).
\item We refined and improved the formalisation of several aspects: substitution, sorts. We also unified some notations (e.g. use of indices in the relations).

 In particular, we improved and specified correctly the nature of the bisimulation relation (FH-bisimulation), its signature, and the impact of the symbolic nature of the relation.

\item We added an example illustrating the difference between our weak bisimulation and weak bisimulation for CCS (more specifically for a process algebra with a choice operator and a $\tau$ action)

\item We improved related works with a positioning relatively to ``negative premisses'' and ``conditional transitions''.

\item The last section has also been extended with a discussion on refinement checking and on tool support.
\end{itemize}

### 2021-10-12 2016-Weak-FH-Bisim/WeakBisim.tex
choice and major points in letter
bullet in letter 2 reviewers

  \begin{itemize}
  \item The sort of a pNet is its signature, i.e. the set of actions in $\AlgAS$ it can
perform, where each action signature is an action 
label plus the arity of the action. In the definition of sorts, we do not need to 
distinguish
input variables and we remove the
\emph{input marker} (?) of variables.\TODO{??? ceci ne se retrouve pas dans la def ci-dessous, c'est une vieille version ?}
\[
\begin{array}{l}
\Sortop(\mylangle S,s_0,V, \to\myrangle) = \{\Sortop(\alpha)|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
    J}\rangle} s'\in \to \} \\
\Sortop(\mylangle \set{\pNet}\!, %\pNet_i^{i\in I}, 
\set{\Sort},
\set{\symb{SV}\,}\myrangle)
=\{\Sortop(\alpha') |\, \SV{\set{\alpha}}{\alpha'}{e_b}\in\set{\symb{SV}}\,\}\\
\Sortop(\alpha(p_1,..,p_n))=(\alpha,n)
\end{array}
\]

@@

A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
is said to be \emph{open}.
Sort comes naturally with a compatibility relation that is similar to a type-compatibility check. We simply say that two sorts are compatible if they consist of the same actions with the same arity. In practice, it is sufficient to check the equality of the two sets of action signatures of the two sorts\footnote{A more complex compatibility relation could be defined, but this is out of the scope of this article.}.
\TODO{Eric: j'ai 2 pbs ici, d'une part, on ne l'utilise que dans la composition de pNet ci-dessous, ou il nous faudrait plutot l'inclusion des sorts, que l'egalite. D'autre part on la definie (Definition 3) comme un ensemble de signatures, mais on cite a plusieurs endroits comme un ensemble d'actions parametrees... }
\end{definition}

@@

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\,\myrangle$
 can be (partially) filled by providing  a pNet $\pNetQ$ to fill one of  its holes.	
	Suppose $j_0\in J$ and \ERIC{$\Sortop(\pNetQ) \subset \Sort_{j_0}$}, then: % are compatible:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNetQ\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNetQ\},\Sort_j^{j\in J\setminus \{j_0\}},
	\set{\symb{SV}}\,\myrangle
	\]
\end{definition}

@@

\begin{example}[CCS choice (counter-example)]\label{ex:CCSplus}
Here is the encoding of a choice operator.\\[1.5ex]
\begin{tabular}{cp{.52\linewidth}}
 \begin{minipage}{.46\linewidth}
 \includegraphics[width=\linewidth]{XFIG/Choice.pdf}\vspace{-18ex}
\end{minipage} 
&
The left hole is
  indexed $l$ the right hole $r$. The third subnet, contains an LTS encoding
  the control part. We obtain the specific behaviour with the 
  synchronisation vector. The first action  of one of
  the holes decides which branch of the LTS is activated;
   all subsequent actions will be from the same side. 
\\
\end{tabular}

>>>>

  \begin{itemize}
  \item The sort of a pNet is its signature, i.e. the set of actions in $\AlgAS$ it can
perform, where each action signature is an action 
label plus the arity of the action.
% Note that, in this definition of sorts we do not  
%distinguish
%input variables from other expressions.
%and only  and we remove the
%\emph{input marker} (?) of variables.\TODO{??? ceci ne se retrouve pas dans la def ci-dessous, c'est une vieille version ?}
\[
\begin{array}{l}
\Sortop(\mylangle S,s_0,V, \to\myrangle) = \{\Sortop(\alpha)|s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
    J}\rangle} s'\in \to \} \\
\Sortop(\mylangle \set{\pNet}\!, %\pNet_i^{i\in I}, 
\set{\Sort},
\set{\symb{SV}\,}\myrangle)
=\{\Sortop(\alpha') |\, \SV{\set{\alpha}}{\alpha'}{e_b}\in\set{\symb{SV}}\,\}\\
\Sortop(\alpha(p_1,..,p_n))=(\alpha,n)
\end{array}
\]

@@

A pNet $Q$ is \emph{closed} if it has no hole: $\Holes(Q)=\emptyset$; else it
is said to be \emph{open}.
Sort comes naturally with a compatibility relation that is similar to a type-compatibility check. We simply say that two sorts are compatible if they consist of the same actions with the same arity. In practice, it is sufficient to check the equality of the two sets of action signatures of the two sorts\footnote{A more complex compatibility relation could be defined, but this is out of the scope of this article.}.
%\TODO{Eric: j'ai 2 pbs ici, d'une part, on ne l'utilise que dans la composition de pNet ci-dessous, ou il nous faudrait plutot l'inclusion des sorts, que l'egalite. D'autre part on la definie (Definition 3) comme un ensemble de signatures, mais on cite a plusieurs endroits comme un ensemble d'actions parametrees... }
\end{definition}

@@

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, 
	\set{\symb{SV}}\,\myrangle$
 can be (partially) filled by providing  a pNet $\pNetQ$ to fill one of  its holes.	
	Suppose $j_0\in J$ and \ERIC{$\Sortop(\pNetQ) \subseteq \Sort_{j_0}$}, then: % are compatible:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNetQ\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNetQ\},\Sort_j^{j\in J\setminus \{j_0\}},
	\set{\symb{SV}}\,\myrangle
	\]
\end{definition}

@@

\begin{example}[CCS choice (counter-example)]\label{ex:CCSplus}
Here is the encoding of a choice operator.\\[1.5ex]
\begin{tabular}{cp{.53\linewidth}}
 \begin{minipage}{.43\linewidth}
 \includegraphics[width=\linewidth]{XFIG/Choice.pdf}\vspace{-18ex}
\end{minipage} 
&
The left hole is
  indexed $l$ the right hole $r$. The third subnet, contains an LTS encoding
  the control part. We obtain the specific behaviour with the 
  synchronisation vector. The first action  of one of
  the holes decides which branch of the LTS is activated;
   all subsequent actions will be from the same side. 
\\
\end{tabular}
